Tutorial_3b

### 2024-10-24 Tutorial_1a.Rmd
update tut1a
add content list to tutorials

# Load the libraries that we will be using.

@@

# Read in the raw data from Panis and Schmidt (2016) - can be skipped.

@@

# Do some wrangling to get a simpler data set - can be skipped.

@@

# Read in simpler data file, change variable names, and create factors.

@@

# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).

@@

# Extract info, save plot for each subject, save a lifetable, create and save the person-trial-bin oriented data file for hazard modeling, and a data file for ca(t) modeling.

@@

@@

Create input file for hazard models.

@@

Create input file for ca(t) models.

@@

# Optional: Create single plot for data aggregated across subjects. This is only recommended when all subjects show the same qualitative patterns.

>>>>

Content:
1. Load libraries
2. Read and wrangle data
3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
4. Extract info (life table, plots, ...) and create data files for fitting hazard and conditional accuracy models
5. Optional: set up life tables and make plots for data aggregated across participants

# 1. Load the libraries that we will be using.

@@

# 2. Read and wrangle data

Read in the raw data from Panis and Schmidt (2016) - can be skipped.

@@

Do some wrangling to get a simpler data set - can be skipped.

@@

Read in simpler data file, change variable names, and create factors.

@@

# 3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)

@@

Save the calculations.

```{r save-descr, eval=F}
save(data_final, file="Tutorial_1_descriptive_stats/data/descr_stats_tut1a.Rdata")
```

# 4. Extract info (life table, plots, ...) and create data files for fitting hazard and conditional accuracy models

@@

Save the plots.

@@

Create input file for fitting hazard models.

@@

Create input file for fitting ca(t) models.

@@

# 5. Optional: set up life tables and make plots for data aggregated across participants

Note: This is only recommended when all subjects show the same qualitative patterns.

### 2024-10-24 Tutorial_1b.Rmd
add content list to tutorials

Tutorial_1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).

# Load the libraries that we will be using.

@@

# Read in the raw data from Panis and Schmidt (2016).

@@

# Do some wrangling.

@@

# Change variable names, and create factors.

@@

# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), P(t), and ca(t)).

@@

# Extract info, and save plot for each subject

>>>>

Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).

Content:
1. Load libraries
2. Read and wrangle data
3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)
4. Extract info (life table, plots, ...)

# 1. Load the libraries that we will be using.

@@

# 2. Read and wrangle data

Read in the raw data from Panis and Schmidt (2016).

@@

Do some wrangling.

@@

Change variable names, and create factors.

@@

# 3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)

@@

# 4. Extract info (life table, plots, ...)

### 2024-10-24 Tutorial_2a.Rmd
add content list to tutorials

# Load the libraries that we will be using

@@

# Set options.

@@

# Load the person-trial-bin data set that we saved in Tutorial 1a.

@@

# Prepare the data set.

@@

# Plot the logit and complementary log-log (cloglog) link functions

@@

# Visualize different prior distributions on the logit and cloglog scales

@@

## Simulate, convert, and fit

@@

# generate draws from U(0,1) and convert
dat <- 
  tibble(p = runif(1e6, 0, 1)) %>% 
  mutate(g = logit(p),
         c = cloglog(p)) 
# display
dat %>%   
  ggplot(aes(x = c)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

plot(density(dat$c))
```

```{r fit-skew-normal, eval=F}
# fit models
fit_skewN <-
  brm(data = dat,
      family = skew_normal(),
      c ~ 1,
      chains = 4, cores = 4,
      file = "Tutorial_2_Bayesian/models/fit_skewN")
```

@@

## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale

@@

# Fit 4 cloglog-hazard models.

@@

### Fit model M1

@@

### Prepare data for M2

@@

### Set up priors for M2

@@

### Fit model M2

@@

### Prepare data for M3

@@

### Set up priors for M3

@@

### Fit model M3

@@

### Prepare data for M4

@@

### Set up priors for M4

@@

### Fit model M4

@@

# Compare 4 models

@@

Compare all three models:

@@

# Plot effects for selected model M4

## Pareto k estimates

@@

## calculate point and interval estimates, and hazard ratios

@@

## Plot model-based predicted cloglog-hazard and hazard functions, for a single subject

@@

### Set up new data

@@

### First, on the cloglog-scale:

@@

### Second, on hazard scale:

>>>>

Content:
1. Load libraries.
2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
3. Plot the logit and complementary log-log (cloglog) link functions.
4. Visualize different prior distributions on the logit and cloglog scales.
5. Fit hazard models.
6. Compare models using loo and waic.
7. Display effects for the selected model.

# 1. Load the libraries that we will be using.

@@

Set options.

@@

# 2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.

@@

Prepare the data set.

@@

# 3. Plot the logit and complementary log-log (cloglog) link functions.

@@

# 4. Visualize different prior distributions on the logit and cloglog scales.

@@

## Simulate, convert, and fit.

@@

# generate draws from U(0,1) and convert
dat <- 
  tibble(p = runif(1e6, 0, 1)) %>% 
  mutate(g = logit(p),
         c = cloglog(p)) 
# display
dat %>%   
  ggplot(aes(x = c)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())
```

```{r fit-skew-normal, eval=F}
# fit model
fit_skewN <-
  brm(data = dat,
      family = skew_normal(),
      c ~ 1,
      chains = 4, cores = 4,
      file = "Tutorial_2_Bayesian/models/fit_skewN")
```

@@

## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale.

@@

# 5. Fit hazard models.

@@

### Fit model M1.

@@

### Prepare data for M2.

@@

### Set up priors for M2.

@@

### Fit model M2.

@@

### Prepare data for M3.

@@

### Set up priors for M3.

@@

### Fit model M3.

@@

### Prepare data for M4.

@@

### Set up priors for M4.

@@

### Fit model M4.

@@

# 6. Compare models using loo and waic.

@@

Compare all three models.

@@

# 7. Display effects for the selected model (model_M4).

## Pareto k estimates.

@@

## calculate point and interval estimates, and hazard ratios.

@@

## Plot model-based predicted cloglog-hazard and hazard functions, for a single subject.

@@

### Set up new data.

@@

### First, on the cloglog-scale.

@@

### Second, on hazard scale.

### 2024-10-24 Tutorial_2b.Rmd
add content list to tutorials

# Load the libraries that we will be using

@@

# Load the person-trial data set that we saved in Tutorial 1a.

@@

# Fit model  M4_ca: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.

## Prepare data for M4_ca

@@

### Set up priors for M4_ca

@@

### Fit model M4_ca

@@

# Display effects

@@

## calculate point and interval estimates, and hazard ratios

@@

## Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject

@@

### First, on the logit-scale:

@@

### Second, on ca(t) scale:

>>>>

Content:
1. Load libraries.
2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
3. Fit a conditional accuracy model.
4. Display effects for the conditional accuracy model.

# 1. Load the libraries that we will be using.

@@

# 2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.

@@

# 3. Fit a conditional accuracy model.

We fit a single model "M4_ca": We use a third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.

## Prepare data for M4_ca.

@@

### Set up priors for M4_ca.

@@

### Fit model M4_ca.

@@

# 4. Display effects for the conditional accuracy model.

@@

## calculate point and interval estimates, and odds ratios.

@@

## Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject.

@@

### First, on the logit-ca(t) scale.

@@

### Second, on the ca(t) scale.

### 2024-10-24 Tutorial_3a.Rmd
add content list to tutorials

# Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")

@@

# Load the person-trial-bin data set that we saved in Tutorial 1.

@@

# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)

@@

@@

Fixed effects:

```{r effects-M3-f}
load("Tutorial_3_Frequentist/model_M3_f.RData")
fixef(model_M3_f)
```

# Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3

@@

```{r extract-estimates}
est <- tibble(brm = fixef(model_M3)[,1],
              glmer = fixef(model_M3_f),
              effects = names(fixef(model_M3_f))) 
```

@@

p1 <- ggplot(data = est, aes(x = effects))
p1 <- p1 + geom_point(aes( y = brm,  color = "c1", shape = "s1"))
p1 <- p1 + geom_point(aes( y = glmer, color = "c2", shape = "s2")) 
p1 <- p1 + labs( x = "Effects", y = "cloglog-hazard" )
p1 <- p1 + scale_color_manual(name = "Model", 
                                breaks = c("c1", "c2"), 
                                values = cols,
                                labels = c("brm", "glmer"))
p1 <- p1 + scale_shape_manual(name = "Model", 
                              breaks = c("s1", "s2"),
                              values = shapes,
                              labels = c("brm", "glmer")) +
   theme(panel.grid = element_blank(),
         panel.background = element_rect(fill="white",color="black"),
         legend.position = "top",
         axis.text.x = element_text(angle=90)) 
p1
```

>>>>

Content:
1. Load libraries.
2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
3. Fit hazard model in the frequentist framework (model_M3_f).
4. Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3.

# 1. Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")

@@

# 2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.

@@

# 3. 3. Fit hazard model in the frequentist framework (model_M3_f)

@@

Model_M3_f contains the same effects as Model_M3 in Tutorial 2a.

@@

```{r effects-M3-f}
load("Tutorial_3_Frequentist/model_M3_f.RData")
summary(model_M3_f)
```

# 4. Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3.

@@

```{r extract-estimates}
est <- tibble(brm = fixef(model_M3)[,c(1,3,4)],
              glmer = tidy(model_M3_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(2,3,7,8)], 
              effectname = names(fixef(model_M3_f)),
              position = 1:16) 
```

@@

p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  color = "c1", shape = "s1")) +
 geom_point(aes( y = glmer$estimate, color = "c2", shape = "s2"), position = position_nudge(x = .1))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], color = "c1"), show.legend = T) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, color = "c2"), show.legend = T, position = position_nudge(x = .1)) +
 geom_hline(yintercept=0,colour = "grey", linetype = 2) +
 scale_x_continuous(breaks = 1:16,labels = est$effectname) +
 labs( x = "Effects in model M3", y = "cloglog-hazard" ) +
 scale_color_manual(name = "Model", 
                                breaks = c("c1", "c2"), 
                                values = cols,
                                labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                              breaks = c("s1", "s2"),
                              values = shapes,
                              labels = c("brm", "glmer")) +
   theme(panel.grid = element_blank(),
         panel.background = element_rect(fill="white",color="black"),
         legend.position = "top",
         axis.text.x = element_text(size=11,angle=90),
         axis.text.y = element_text(size=13),
         axis.title.x = element_text(size = 14),
         axis.title.y = element_text(size = 14),
         legend.text=element_text(size=13)) 
p1
```

### 2024-10-24 Tutorial_3b.Rmd
add fixed effects code
add content list to tutorials

# Load the libraries that we will be using

@@

# Load the data set for ca(t) modeling that we saved in Tutorial 1a.

@@

# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)

@@

@@

>>>>

Content:
1. Load libraries.
2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.
3. Fit a conditional accuracy model in the frequentist framework (model M3_ca_f).

# 1. Load the libraries that we will be using.

@@

# 2. Load and wrangle the person-trial data set that we saved in Tutorial 1a.

@@

# 3. Fit a conditional accuracy model in the frequentist framework (model M3_ca_f).

@@

Model_M3_ca_f contains the same effects as Model_M3 in Tutorial 2a.

@@

```{r effects-M3-f}
load("Tutorial_3_Frequentist/model_M3_ca_f.RData")
summary(model_M3_ca_f)
```
